Discrete Quantum Theories and Computing

### 2019-01-30 dissertation.tex
dissertation.lyx Resolving TODOs before "5.1.2 States Consistent with Classical and Quantum IVPMs"...

In summary, the full space of a $D$-dimensional quantum state, including
its overall phase defining its relationship to other quantum states,
is the topological space $\Sphere{2D-1}$. For an isolated system,
the overall phase is not measurable, and eliminating the phase dependence
corresponds to identifying $\Sphere{2D-1}$ as a circle bundle over
the base space $\CP{D-1}$, and therefore $\CP{D-1}$ defines the
$2D-2$ intrinsic, irreducible, degrees of freedom of the isolated
$D$-dimensional state's dynamics. In mathematical notation, this
would be written $\Sphere{1}\hookrightarrow\Sphere{2D-1}\rightarrow\CP{D-1}$.
\todo{Citation? }For $D=2$, the single qubit, we have $2-1=1$,
and the base space of the circle bundle is $\CP{1}=\Sphere{2}$, the
usual Bloch sphere. Note that only for $D=2$ is this actually a sphere-like
geometry due to an accident of low-dimensional topology.

@@

Given a normalized pure state $\ket{\Psi}=\sum_{i=0}^{D-1}\alpha_{i}\ket{i}$,
a natural quantity characterizing a $D$-dimensional system is its
\textit{density matrix}, $\rho=\op{\Psi}{\Psi}$, or 
\begin{equation}
\rho=\begin{pmatrix}|\alpha_{0}|^{2} & \alpha_{0}\alpha_{1}^{*} & \cdots & \alpha_{0}\alpha_{D-1}^{*}\\
\alpha_{1}\alpha_{0}^{*} & |\alpha_{1}|^{2} & \cdots & \alpha_{1}\alpha_{D-1}^{*}\\
\vdots & \vdots & \ddots & \vdots\\
\alpha_{D-1}\alpha_{0}^{*} & \cdots & \alpha_{D-1}\alpha_{D-2}^{*} & |\alpha_{D-1}|^{2}
\end{pmatrix}\,.\label{nqDensity.eq}
\end{equation}
We can now use the complex generalization of the classical Veronese
coordinate system for projective geometry to remove the overall phase
ambiguity $\rme^{\rmi\theta}$ from the $D$-dimensional states. If
we take a particular weighting of the elements of the density matrix~$\rho$,
we can construct a \textit{unit vector} of real dimension $D^{2}$
with the form: 
\begin{equation}
\Hat{a}=\left(|\alpha_{i}|^{2},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,\label{nqCoords.eq}
\end{equation}
where 
\[
\Hat{a}\cdot\Hat{a}=\sum_{i=0}^{D-1}\left(\left|\alpha_{i}\right|^{2}\right)^{2}+\sum_{i=0}^{D-1}\sum_{\substack{j=0\\
j\ne i
}
}^{D-1}\left(\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}+\left(\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}=\left(\sum_{i=0}^{D-1}\left|\alpha_{i}\right|^{2}\right)\left(\sum_{j=0}^{D-1}\left|\alpha_{j}\right|^{2}\right)=1\,.
\]
This construction gives an explicit embedding of the $\left(D-1\right)$-dimensional
complex, or $\left(2D-2\right)$-dimensional real, object in a real
space of dimension $D^{2}$. However, this is somewhat subtle because
the vector is of unit length, so technically the embedding space is
a sphere of dimension $D^{2}-1$ embedded in $\R^{D^{2}}$. For example,
the two-dimensional irreducible states could be represented in a four-dimensional
embedding, but the magnitude of every coordinate would be one; furthermore,
the object embedded in the resulting $\Sphere{3}$ is indeed $\Sphere{2}$
because we can fix one complex coordinate to be unity, and let one
vary, giving a total of two irreducible dimensions. In fact, one must
choose \textit{two} coordinate patches, one covering one pole of $\Sphere{2}$
with coordinates $\alpha_{0}=1+\rmi0$ and $\alpha_{1}=x_{1}+\rmi y_{1}$,
and the other patch covering the other pole of $\Sphere{2}$ with
coordinates $\alpha_{0}=x_{0}+\rmi y_{0}$ and $\alpha_{1}=1+\rmi0$.
\todo{Explain the last part clearer or add a picture.}

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on both
sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This identification can be achieved by mapping a classical
event~$E$ to the projector generated by $E$,
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$, because
$\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)$ is equal to
$\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)$ as operators,
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ are mapped to
the same projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Therefore, for any classical event~$E$, its corresponding quantum
event is defined to be the projector $\varphi\left(E\right)$, and
the set of all projectors on a given Hilbert space is called a quantum
event space~$\events$. \todo{Check the citation of this part. Did
von Neumann really think about this when he uses projectors to formulate
the quantum theory?}

@@

Given a Hilbert space $\mathcal{H}$ of dimension $D$ and a probability
assignment for every projector $P$, we can define the expectation
value of an observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$, with eigenvalues $\lambda_{i}\in\mathbb{R}$,
as \cite{544199,Jaeger2007}: 
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=0}^{D-1}\lambda_{i}\mu\left(P_{i}\right)\,,\label{eq:quantum-expectation}
\end{equation}
where the subscript $\mu$ might be omitted if it is clear according
to the context. This definition is also consistent with the classical
expectation values because we can pullback an observable to a classical
random variable, and the expectation values are invariant.\todo{
Add interpretation!}

\begin{definition}[Pullback of Observables]\label{def:observable2random-variable}Consider
an observable~$\mathbf{O}$ diagonalizable by an orthonormal basis
$\Omega=\left\{ \ket{0},\ket{1},\ldots,\ket{D-1}\right\} $ so that
$\mathbf{O}$ has spectral decomposition $\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\proj{i}$.
If we restrict $\varphi$ on the classical event space~$2^{\Omega}$
and consider $\varphi\colon2^{\Omega}\rightarrow\events$, then the
pullback of $\mathbf{O}$ by $\varphi$ is a random variable $\varphi^{*}\mathbf{O}\colon\Omega\rightarrow\mathbb{R}$
defined by $\varphi^{*}\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\mathbf{1}_{\left\{ \ket{i}\right\} }$.
\todo{Put the definition of indicator function!}\end{definition}

@@

In general, \todo{Ask Gerardo why? }the reduced density operator~$\rho_{j}$
of the $j$-th subsystem can always be expressed as

@@

Our computations with this ``toy'' modal quantum theory showed that
it possesses ``supernatural'' computational power. For example,
one can solve a black box version of the \usat\ problem~\cite{usat}
in a way that outperforms conventional quantum computing. The classical
\usat\ problem (also known as $\lang{USAT}$ or $\lang{UNAMBIGUOUS-SAT}$)
\todo{Add citation about where these two names come from. }is the
problem of deciding whether a given Boolean formula has a satisfying
assignment, assuming that it has at most one such assignment~\cite{Papadimitriou1993}.
This problem is, in a precise sense~\cite{Valiant198685}, just as
hard as the general satisfiability problem and hence all problems
in the $\NP$ complexity class. Our black-box version of the \usat\ problem
replaces the Boolean formula with an arbitrary black box. Solutions
to this generalized problem can be used to solve an unstructured database
search of size $N$ using $O\left(\log N\right)$ black box evaluations
by binary search on the database. This algorithm then outperforms
the known asymptotic bound $O\left(\sqrt{N}\right)$ for unstructured
database search in conventional quantum computing.

@@

\begin{proof}\todo{This part use $\ket{\overline{a}}=\ket{a_{1}}\ldots\ket{a_{n}}$
while previous parts use $\ket{\Psi}=\ket{\psi_{1}}\cdots\ket{\psi_{j}}\cdots\ket{\psi_{n}}$.
Moreover, the ``bar'' is heavily used in QIVPM discussion later,
so maybe not using bar here???? }The state is initialized to $\ket{0}\ket{\overline{0}}$,
with $\ket{\overline{0}}=\ket{0}\ket{0}\cdots\ket{0}$, i.e., the
tensor product of $n$ $\ket{0}$ states. As Eq.~(\ref{eq:MQT-dynamics-apply}),
applying the map $S$ to each qubit in the second component of the
state produces $\ket{0}\ket{\overline{+}}$ where $\ket{\overline{+}}$
denotes the sequence $\ket{+}\ldots\ket{+}$ of length $n$. Applying
$\uf$ to the entire state has no effect since $U_{f}$ is the identity
when $f$ is unsatisfiable. Applying $S$ to each qubit in the second
component of the state produces $\ket{0}\ket{\overline{0}}$. Applying
$S^{\dagger}$ to the first component leaves the state unchanged.
As the first component of the state is $0$, applying the map $\sigma_{0}$
(which is the identity) leaves the state unchanged. \todo{Control-not
needs to be defined and explained in Sec.~\ref{subsec:Quantum-Circuits}
}Applying $S^{\dagger}$ to the first component leaves the state
unchanged. Measuring the state will deterministically produce $\ket{0}\ket{\overline{0}}$.\end{proof}

@@

\begin{proof}Assume the function $f$ is satisfiable at some input
$a_{1},a_{2},\ldots,a_{n}$ denoted $\overline{a}$, and where $\ket{\overline{a}}=\ket{a_{1}}\ldots\ket{a_{n}}$.
In the second step, the state becomes $\ket{0}\ket{\overline{+}}$
as above. We can write this state as $\ket{0}\ket{\overline{a}}+\Sigma_{\overline{x}\neq\overline{a}}\ket{0}\ket{\overline{x}}$.
Applying $U_{f}$ produces $\ket{1}\ket{\overline{a}}+\Sigma_{\overline{x}\neq\overline{a}}\ket{0}\ket{\overline{x}}$.
We can rewrite this state as $\ket{+}\ket{\overline{a}}+\Sigma_{\overline{x}}\ket{0}\ket{\overline{x}}=\ket{+}\ket{\overline{a}}+\ket{0}\ket{\overline{+}}$,
where the summation is now over all vectors (notice that $\ket{0}\ket{\overline{a}}+\ket{0}\ket{\overline{a}}$
is the zero vector). Applying $S$ to each qubit in the second component
produces $\ket{+}\ket{\overline{S(a)}}+\ket{0}\ket{\overline{0}}$.
Applying $S^{\dagger}$ to the first component produces: $\ket{1}\ket{\overline{S(a)}}+\ket{0}\ket{\overline{0}}$.
Applying control-not gate, which applying $\sigma_{0}$ or $\sigma_{x}$
on the second component depending on the first component of the state,
and produces 
\begin{equation}
\ket{1}\left(\sigma_{x}\ket{\overline{S(a)}}\right)+\ket{0}\left(\sigma_{0}\ket{\overline{0}}\right)=\ket{1}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}\,.
\end{equation}
Applying $S^{\dagger}$ to the first component produces $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}$.
For the measurement of $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}$
to be guaranteed to never be $\ket{0}\ket{\overline{0}}$, we need
to verify that $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}$
has one occurrence $\ket{0}\ket{\overline{0}}$. \todo{The following
need to be rewritten. }This can be easily proved as follows. Since
each $a_{i}$ is either 0 or 1, then each $S(a_{i})$ is either $+$
or $1$, and hence each $\mathsf{not}\left(S(a_{i})\right)$ is either~$+$
or~$0$. The result follows since any state with a combination of
$+$ and $0$, when expressed in the standard basis, would consist
of a superposition containing the state $\ket{0\ldots}$.\end{proof}

@@

We can now compute the size of the equivalence class of complex unit-modulus
phases corresponding to the Hopf fibration circle. \todo{Explain
the Hopf fibration circle here\ldots ? }Since $\Fpp$ has $p^{2}-1$
non-zero values, and the map $\fnorm{\alpha}$ distributes these equally
across the domain of $p-1$ non-zero elements $c\in\Fp$, there are
$\frac{p^{2}-1}{p-1}=p+1$ (non-zero) domain elements in $\Fpp$ for
each (non-zero) image element in $\Fp$. We illustrate this graphically
in Figure~\ref{proof1.fig}. Thus, the Hopf circle always has the
size $p+1$, corresponding essentially to a discrete projective line,
and that is the size of each equivalence class of the map $\fnorm{\alpha}$
for non-vanishing $\alpha$, including the map to the unit norm value
$c=1\in\Fp$.
\begin{figure}
\noindent \centering{}\includegraphics[width=3.6in]{Fpp2Fp}\label{proof1.fig}\caption{Sketch of the map from $\Fpp$ to $\Fp$ using $\fnorm{\alpha}$,
showing the decomposition of $\Fpp$ into the zero element $\left(0,0\right)$
and the $p^{2}-1=\left(p+1\right)\left(p-1\right)$ non-zero elements
that map onto the $p-1$ non-zero elements of $\Fp$ with multiplicity
$p+1$.}
\end{figure}
\end{proof}

@@

This process of describing the discrete $D$-dimensional irreducible
states can again be understood geometrically by following the discrete
analog of the Hopf fibration. First, we construct the discrete version
of the quadratic unit-length form that automatically annihilates the
distinction among states differing only by a discrete phase, 
\begin{equation}
\Hat{a}=\left(\fnorm{\alpha_{i}},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,
\end{equation}
where 
\begin{equation}
\Hat{a}\cdot\Hat{a}=\left(\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}\right)^{2}=1\,.
\end{equation}
From Proposition~\ref{appH.sec}, we know that $p+1$ elements of
this discrete $\Sphere{2D-1}$ structure map to the \textit{same point}
in $\Hat{a}$. \todo{Does Proposition~\ref{appH.sec} really show
this? }Each set of $p+1$ redundant points is, geometrically speaking,
the \textit{discrete Hopf fibration circle} living above each \textit{irreducible}
point of the $D$-dimensional state description. These $p+1$ points
are interpretable as the $p$ finite points plus the single point
at infinity of the projective discrete line (see, e.g.,~\cite{arnold_2010}).

@@

The cardinal probability over finite fields defined in Eq.~(\ref{eq:cardinalProbability})
is based on the conventional Born rule, Eq.~(\ref{eq:Born-ComputationalBasis}).
Although they are valid inside an ordered region, some quantum algorithms,
like Shor's~\cite{Shor1999,Mermin2007,Jaeger2007}, don't localize
their computation in a small region. \todo{Remove the discussion
of Shor's algorithm if I cannot verify it on DQT~(I) on time. }Moreover,
since we haven't found a way to manipulate the resulting cardinal
probabilities, we have little idea how to define the mixed states
and expectation values. To overcome these issues, we try a different
approach to define a ``conventional'' probability over finite fields
axiomatically.

@@

\todo{Add some introductory text\ldots}

>>>>

In summary, the full space of a $D$-dimensional quantum state, including
its overall phase defining its relationship to other quantum states,
is the topological space $\Sphere{2D-1}$. For an isolated system,
the overall phase is not measurable, and eliminating the phase dependence
corresponds to identifying $\Sphere{2D-1}$ as a circle bundle over
the base space $\CP{D-1}$, and therefore $\CP{D-1}$ defines the
$2D-2$ intrinsic, irreducible, degrees of freedom of the isolated
$D$-dimensional state's dynamics. In mathematical notation, this
would be written $\Sphere{1}\hookrightarrow\Sphere{2D-1}\rightarrow\CP{D-1}$.
\begin{comment}
\todo{Citation? }
\end{comment}
For $D=2$, the single qubit, we have $2-1=1$, and the base space
of the circle bundle is $\CP{1}=\Sphere{2}$, the usual Bloch sphere.
Note that only for $D=2$ is this actually a sphere-like geometry
due to an accident of low-dimensional topology.

@@

Given a normalized pure state $\ket{\Psi}=\sum_{i=0}^{D-1}\alpha_{i}\ket{i}$,
a natural quantity characterizing a $D$-dimensional system is its
\textit{density matrix}, $\rho=\op{\Psi}{\Psi}$, or 
\begin{equation}
\rho=\begin{pmatrix}|\alpha_{0}|^{2} & \alpha_{0}\alpha_{1}^{*} & \cdots & \alpha_{0}\alpha_{D-1}^{*}\\
\alpha_{1}\alpha_{0}^{*} & |\alpha_{1}|^{2} & \cdots & \alpha_{1}\alpha_{D-1}^{*}\\
\vdots & \vdots & \ddots & \vdots\\
\alpha_{D-1}\alpha_{0}^{*} & \cdots & \alpha_{D-1}\alpha_{D-2}^{*} & |\alpha_{D-1}|^{2}
\end{pmatrix}\,.\label{nqDensity.eq}
\end{equation}
We can now use the complex generalization of the classical Veronese
coordinate system for projective geometry to remove the overall phase
ambiguity $\rme^{\rmi\theta}$ from the $D$-dimensional states. If
we take a particular weighting of the elements of the density matrix~$\rho$,
we can construct a \textit{unit vector} of real dimension $D^{2}$
with the form: 
\begin{equation}
\Hat{a}=\left(|\alpha_{i}|^{2},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,\label{nqCoords.eq}
\end{equation}
where 
\[
\Hat{a}\cdot\Hat{a}=\sum_{i=0}^{D-1}\left(\left|\alpha_{i}\right|^{2}\right)^{2}+\sum_{i=0}^{D-1}\sum_{\substack{j=0\\
j\ne i
}
}^{D-1}\left(\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}+\left(\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*}\right)^{2}=\left(\sum_{i=0}^{D-1}\left|\alpha_{i}\right|^{2}\right)\left(\sum_{j=0}^{D-1}\left|\alpha_{j}\right|^{2}\right)=1\,.
\]
This construction gives an explicit embedding of the $\left(D-1\right)$-dimensional
complex, or $\left(2D-2\right)$-dimensional real, object in a real
space of dimension $D^{2}$. However, this is somewhat subtle because
the vector is of unit length, so technically the embedding space is
a sphere of dimension $D^{2}-1$ embedded in $\R^{D^{2}}$. For example,
the two-dimensional irreducible states could be represented in a four-dimensional
embedding, but the magnitude of every coordinate would be one; furthermore,
the object embedded in the resulting $\Sphere{3}$ is indeed $\Sphere{2}$
because we can fix one complex coordinate to be unity, and let one
vary, giving a total of two irreducible dimensions. In fact, one must
choose \textit{two} coordinate patches, one covering one pole of $\Sphere{2}$
with coordinates $\alpha_{0}=1+\rmi0$ and $\alpha_{1}=x_{1}+\rmi y_{1}$,
and the other patch covering the other pole of $\Sphere{2}$ with
coordinates $\alpha_{0}=x_{0}+\rmi y_{0}$ and $\alpha_{1}=1+\rmi0$.
\begin{comment}
\todo{Explain the last part clearer or add a picture.}
\end{comment}

@@

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on both
sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This identification can be achieved by mapping a classical
event~$E$ to the projector generated by $E$,
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$, because
$\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)$ is equal to
$\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)$ as operators,
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ are mapped to
the same projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Therefore, for any classical event~$E$, its corresponding quantum
event is defined to be the projector $\varphi\left(E\right)$, and
the set of all projectors on a given Hilbert space is called a quantum
event space~$\events$. %
\begin{comment}
\todo{Check the citation of this part. Did von Neumann really think
about this when he uses projectors to formulate the quantum theory?}
\end{comment}

@@

Given a Hilbert space $\mathcal{H}$ of dimension $D$ and a probability
assignment for every projector $P$, we can define the expectation
value of an observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$, with eigenvalues $\lambda_{i}\in\mathbb{R}$,
as \cite{544199,Jaeger2007}: 
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=0}^{D-1}\lambda_{i}\mu\left(P_{i}\right)\,,\label{eq:quantum-expectation}
\end{equation}
where the subscript $\mu$ might be omitted if it is clear according
to the context. This definition is also consistent with the classical
expectation values because we can pullback an observable to a classical
random variable, and the expectation values are invariant.%
\begin{comment}
\todo{ Add interpretation!}
\end{comment}

\begin{definition}[Pullback of Observables]\label{def:observable2random-variable}Consider
an observable~$\mathbf{O}$ diagonalizable by an orthonormal basis
$\Omega=\left\{ \ket{0},\ket{1},\ldots,\ket{D-1}\right\} $ so that
$\mathbf{O}$ has spectral decomposition $\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\proj{i}$.
If we restrict $\varphi$ on the classical event space~$2^{\Omega}$
and consider $\varphi\colon2^{\Omega}\rightarrow\events$, then the
pullback of $\mathbf{O}$ by $\varphi$ is a random variable $\varphi^{*}\mathbf{O}\colon\Omega\rightarrow\mathbb{R}$
defined by $\varphi^{*}\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\mathbf{1}_{\left\{ \ket{i}\right\} }$,
where $\mathbf{1}_{E}$ is the indicator function defined by
\begin{equation}
\mathbf{1}_{E}\left(\omega\right)=\begin{cases}
1 & \textrm{if }\omega\in E\,;\\
0 & \textrm{if }\omega\notin E\,.
\end{cases}
\end{equation}
\begin{comment}
\todo{Put the definition of indicator function!}
\end{comment}
\end{definition}

@@

In general, %
\todo{Ask Gerardo why? }
the reduced density operator~$\rho_{j}$ of the $j$-th subsystem
can always be expressed as

@@

Our computations with this ``toy'' modal quantum theory showed that
it possesses ``supernatural'' computational power. For example,
one can solve a black box version of the \usat\ problem~\cite{usat}
in a way that outperforms conventional quantum computing. The classical
\usat\ problem (also known as $\lang{USAT}$ or $\lang{UNAMBIGUOUS-SAT}$)
\begin{comment}
\todo{Add citation about where these two names come from. }
\end{comment}
is the problem of deciding whether a given Boolean formula has a satisfying
assignment, assuming that it has at most one such assignment~\cite{Papadimitriou1993}.
This problem is, in a precise sense~\cite{Valiant198685}, just as
hard as the general satisfiability problem and hence all problems
in the $\NP$ complexity class. Our black-box version of the \usat\ problem
replaces the Boolean formula with an arbitrary black box. Solutions
to this generalized problem can be used to solve an unstructured database
search of size $N$ using $O\left(\log N\right)$ black box evaluations
by binary search on the database. This algorithm then outperforms
the known asymptotic bound $O\left(\sqrt{N}\right)$ for unstructured
database search in conventional quantum computing.

@@

\begin{proof}%
\begin{comment}
\todo{This part use $\ket{\overline{a}}=\ket{a_{1}}\ldots\ket{a_{n}}$
while previous parts use $\ket{\Psi}=\ket{\psi_{1}}\cdots\ket{\psi_{j}}\cdots\ket{\psi_{n}}$.
Moreover, the ``bar'' is heavily used in QIVPM discussion later,
so maybe not using bar here???? }
\end{comment}
The state is initialized to $\ket{0}\ket{\overline{0}}$, with $\ket{\overline{0}}=\ket{0}\ket{0}\cdots\ket{0}$,
i.e., the tensor product of $n$ $\ket{0}$ states. As Eq.~(\ref{eq:MQT-dynamics-apply}),
applying the map $S$ to each qubit in the second component of the
state produces $\ket{0}\ket{\overline{+}}$ where $\ket{\overline{+}}$
denotes the sequence $\ket{+}\ldots\ket{+}$ of length $n$. Applying
$\uf$ to the entire state has no effect since $U_{f}$ is the identity
when $f$ is unsatisfiable. Applying $S$ to each qubit in the second
component of the state produces $\ket{0}\ket{\overline{0}}$. Applying
$S^{\dagger}$ to the first component leaves the state unchanged.
As the first component of the state is $0$, applying the map $\sigma_{0}$
(which is the identity) leaves the state unchanged. %
\begin{comment}
\todo{Control-not needs to be defined and explained in Sec.~\ref{subsec:Quantum-Circuits}
}
\end{comment}
Applying $S^{\dagger}$ to the first component leaves the state unchanged.
Measuring the state will deterministically produce $\ket{0}\ket{\overline{0}}$.\end{proof}

@@

\begin{proof}Assume the function $f$ is satisfiable at some input
$a_{1},a_{2},\ldots,a_{n}$ denoted $\overline{a}$, and where $\ket{\overline{a}}=\ket{a_{1}}\ldots\ket{a_{n}}$.
In the second step, the state becomes $\ket{0}\ket{\overline{+}}$
as above. We can write this state as $\ket{0}\ket{\overline{a}}+\Sigma_{\overline{x}\neq\overline{a}}\ket{0}\ket{\overline{x}}$.
Applying $U_{f}$ produces $\ket{1}\ket{\overline{a}}+\Sigma_{\overline{x}\neq\overline{a}}\ket{0}\ket{\overline{x}}$.
We can rewrite this state as $\ket{+}\ket{\overline{a}}+\Sigma_{\overline{x}}\ket{0}\ket{\overline{x}}=\ket{+}\ket{\overline{a}}+\ket{0}\ket{\overline{+}}$,
where the summation is now over all vectors (notice that $\ket{0}\ket{\overline{a}}+\ket{0}\ket{\overline{a}}$
is the zero vector). Applying $S$ to each qubit in the second component
produces $\ket{+}\ket{\overline{S(a)}}+\ket{0}\ket{\overline{0}}$.
Applying $S^{\dagger}$ to the first component produces: $\ket{1}\ket{\overline{S(a)}}+\ket{0}\ket{\overline{0}}$.
Applying control-not gate, which applying $\sigma_{0}$ or $\sigma_{x}$
on the second component depending on the first component of the state,
and produces 
\begin{equation}
\ket{1}\left(\sigma_{x}\ket{\overline{S(a)}}\right)+\ket{0}\left(\sigma_{0}\ket{\overline{0}}\right)=\ket{1}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}\,.
\end{equation}
Applying $S^{\dagger}$ to the first component produces $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}$.
For the measurement of $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}+\ket{0}\ket{\overline{0}}$
to be guaranteed to never be $\ket{0}\ket{\overline{0}}$, we need
to verify that $\ket{+}\ket{\overline{\mathsf{not}\left(S(a)\right)}}$
has one occurrence $\ket{0}\ket{\overline{0}}$. %
\begin{comment}
\todo{The following need to be rewritten. }
\end{comment}
This can be easily proved as follows. Since each $a_{i}$ is either
0 or 1, then each $S(a_{i})$ is either $+$ or $1$, and hence each
$\mathsf{not}\left(S(a_{i})\right)$ is either~$+$ or~$0$. The
result follows since any state with a combination of $+$ and $0$,
when expressed in the standard basis, would consist of a superposition
containing the state $\ket{0\ldots}$.\end{proof}

@@

We can now compute the size of the equivalence class of complex unit-modulus
phases corresponding to the Hopf fibration circle. %
\begin{comment}
\todo{Explain the Hopf fibration circle here\ldots ? }
\end{comment}
Since $\Fpp$ has $p^{2}-1$ non-zero values, and the map $\fnorm{\alpha}$
distributes these equally across the domain of $p-1$ non-zero elements
$c\in\Fp$, there are $\frac{p^{2}-1}{p-1}=p+1$ (non-zero) domain
elements in $\Fpp$ for each (non-zero) image element in $\Fp$. We
illustrate this graphically in Figure~\ref{proof1.fig}. Thus, the
Hopf circle always has the size $p+1$, corresponding essentially
to a discrete projective line, and that is the size of each equivalence
class of the map $\fnorm{\alpha}$ for non-vanishing $\alpha$, including
the map to the unit norm value $c=1\in\Fp$.
\begin{figure}
\noindent \centering{}\includegraphics[width=3.6in]{Fpp2Fp}\label{proof1.fig}\caption{Sketch of the map from $\Fpp$ to $\Fp$ using $\fnorm{\alpha}$,
showing the decomposition of $\Fpp$ into the zero element $\left(0,0\right)$
and the $p^{2}-1=\left(p+1\right)\left(p-1\right)$ non-zero elements
that map onto the $p-1$ non-zero elements of $\Fp$ with multiplicity
$p+1$.}
\end{figure}
\end{proof}

@@

This process of describing the discrete $D$-dimensional irreducible
states can again be understood geometrically by following the discrete
analog of the Hopf fibration. First, we construct the discrete version
of the quadratic unit-length form that automatically annihilates the
distinction among states differing only by a discrete phase, 
\begin{equation}
\Hat{a}=\left(\fnorm{\alpha_{i}},\ldots,\sqrt{2}\,\mathrm{Re}\ \alpha_{i}\alpha_{j}^{*},\ldots,\sqrt{2}\,\mathrm{Im}\ \alpha_{i}\alpha_{j}^{*},\ldots\right)\,,
\end{equation}
where 
\begin{equation}
\Hat{a}\cdot\Hat{a}=\left(\sum_{i=0}^{D-1}\fnorm{\alpha_{i}}\right)^{2}=1\,.
\end{equation}
From Proposition~\ref{appH.sec}, we know that $p+1$ elements of
this discrete $\Sphere{2D-1}$ structure map to the \textit{same point}
in $\Hat{a}$. %
\begin{comment}
\todo{Does Proposition~\ref{appH.sec} really show this? }
\end{comment}
Each set of $p+1$ redundant points is, geometrically speaking, the
\textit{discrete Hopf fibration circle} living above each \textit{irreducible}
point of the $D$-dimensional state description. These $p+1$ points
are interpretable as the $p$ finite points plus the single point
at infinity of the projective discrete line (see, e.g.,~\cite{arnold_2010}).

@@

The cardinal probability over finite fields defined in Eq.~(\ref{eq:cardinalProbability})
is based on the conventional Born rule, Eq.~(\ref{eq:Born-ComputationalBasis}).
\begin{comment}
Although they are valid inside an ordered region, some quantum algorithms,
like Shor's~\cite{Shor1999,Mermin2007,Jaeger2007}, don't localize
their computation in a small region. \todo{Remove the discussion
of Shor's algorithm if I cannot verify it on DQT~(I) on time. }Moreover, 
\end{comment}
Since we haven't found a way to manipulate the resulting cardinal
probabilities, we have little idea how to define the mixed states
and expectation values. To overcome these issues, we try a different
approach to define a ``conventional'' probability over finite fields
axiomatically.

@@

\begin{comment}
\todo{Add some introductory text\ldots}
\end{comment}
